 Austral Spring Extratropical Southern Hemisphere zonally assymetric circulation

### 2021-07-26 analysis/paper/shceof.Rmd
Agregrega amplitud de ENSO en conclusiones
Closes #53

```{r enso-phase, fig.cap = " ONI plotted against cEOF2 phase for th 1979 -- 2019 period. Years with magnitude of cEOF2 greater or smaller than the 50th percentile are shown as orange circles and green diamomnds respectively. Black line is the fit ONI ~ sin(phase) computed by weighted OLS using the magnitude of the cEOF2 as weights.", fig.width = width_column, fig.height=3}
gdata <- ceof[, eof[[1]]$left, by = .(season)] %>%
  .[cEOF == "cEOF2"] %>% 
  # .[, hgt := Anomaly(hgt)] %>% 
  .[, arg := Arg(hgt)] %>%
  .[, mag := Mod(hgt)] %>%
  .[enso, on = "time"] %>%
  na.omit()

@@

\textcolor{red}{¿Qué significa para el fit "weighted by the magnitude of cEOF2"?}

@@

\textcolor{red}{Se que la cantidad de años hace bastante limitada la muestra, pero creo que es un resultado a tener en cuenta que en esos eventos muy intensos se alineen. ¿Probaste hacer la correlación con todos los puntos naranjas?}

@@

```{r sst-psi-1, fig.cap = "(ref:sst-psi-1-cap)", fig.width = 7, fig.height=4, dependson="plot_sst_psi", fig.env = "figure*"}
plot_sst_psi(which_cEOF = "cEOF1", which_lev = 200) +
  scale_fill_divergent_discretised("SST",
                                   limits = c(-.3, .3),
                                   guide = guide_colorsteps_bottom(20, order = 2)) 
```

@@

THe second complex EOF represents a wave-3 pattern with maximum magnitude in the Pacific sector.
Essentially, it is an alternate representation of the PSA1 and PSA2 patterns [@mo2001].
We show that the Imaginary cEOF2 can be identified with the PSA1 and the Real cEOF2 with the PSA2.
This cEOF prefers a phase aligned with the Imaginary cEOF2.
This is because the when ENSO is active, it forces the cEOF2 to be in the Imaginary phase.
When ENSO is neutral, the cEOF2 is still active, but with no particular phase.
This mirrors the results of @cai2002, who showed that the CSIRO Model can create PSA-like variability even in the absence of ENSO forcing (with a climatological run), but the variability of one of the PSA modes was enhanced when adding the ENSO signal.

>>>>

```{r enso-phase, fig.cap = "ONI plotted against cEOF2 phase for th 1979 -- 2019 period. Years with magnitude of cEOF2 greater or smaller than the 50th percentile are shown as orange circles and green diamomnds respectively. Black line is the fit ONI ~ sin(phase) computed by weighted OLS using the magnitude of the cEOF2 as weights.", fig.width = width_column, fig.height=3}
gdata <- ceof[, eof[[1]]$left, by = .(season)] %>%
  .[cEOF == "cEOF2"] %>% 
  # .[, hgt := Anomaly(hgt)] %>% 
  .[, arg := Arg(hgt)] %>%
  .[, mag := Mod(hgt)] %>%
  .[enso, on = "time"] %>%
  na.omit()

@@

@@

@@

```{r sst-psi-1, fig.cap = "(ref:sst-psi-1-cap)", fig.width = 7, fig.height=4, dependson="plot_sst_psi", fig.env = "figure*"}
plot_sst_psi(which_cEOF = "cEOF1", which_lev = 200) +
  scale_fill_divergent_discretised("SST",
                                   limits = c(-1, 1),
                                   guide = guide_colorsteps_bottom(20, order = 2)) 
```

@@

The second complex EOF represents a wave-3 pattern with maximum magnitude in the Pacific sector.
Essentially, it is an alternate representation of the PSA1 and PSA2 patterns [@mo2001].
We show that the Imaginary cEOF2 can be identified with the PSA1 and the Real cEOF2 with the PSA2.
This cEOF prefers a phase aligned with the Imaginary cEOF2.
This is because the when ENSO is active, it forces the cEOF2 to be in the Imaginary phase.
When ENSO is neutral, the cEOF2 is still active, but with no particular preferred phase.
This mirrors the results of @cai2002, who showed that the CSIRO Model can create PSA-like variability even in the absence of ENSO forcing (with a climatological run), but the variability of one of the PSA modes was enhanced when adding the ENSO signal. The magnitude of the cEOF2 appears not to be related to the magnitude of ENSO, except frot the three strongest ENSO events in the period (`r knitr::combine_words(year(gdata[order(-oni)][1:3]$time))`) which coincide with the three strongest cEOF2 years.
